Intrinsic alignments from projected tidal fields: non-linear impact on cosmic shear probes

### 2024-04-30 sections/2_cosmic_shear_stats.tex
Updates from Overleaf

Shear two-point correlation functions (\gtwopcf hereafter) can be predicted with percent-level precision and are therefore an ideal quantity to validate weak lensing simulations.
In the Limber approximation\footnote{See \citet{Kilbinger17} for a comparison between the Limber approximation and the exact calculations.}, the lensing angular power spectrum $C^{ij}_{\ell}$ (in spherical harmonics space) in tomographic bins $i$ and $j$ is obtained from an integral over the three-dimensional matter power spectrum $P_{\delta}(k, z)$ (in Fourier space, and redshift $z$ dependant):
  \begin{eqnarray}
C_{\ell}^{ij} = \int_0^{\chi_{\rm H}}  \frac{q^i(\chi) \,q^j(\chi)}{\chi^2} \, P_{\delta}\, \bigg(\frac{\ell+1/2}{\chi},z(\chi)\bigg) \ {\rm d}\chi,
\label{eq:C_ell}
\end{eqnarray}
where $\ell$ is a multipole, $\chi_{\rm H}$ is the co-moving distance to the horizon, $c$ is the speed of light, $H_0$ the Hubble parameter, and the lensing kernels $q^{i}$ and $q^j$ are given by:
\begin{eqnarray}
q^{i(j)}(\chi) = \frac{3}{2}\Omega_{\rm m} \, \bigg(\frac{H_0}{c} \bigg)^2 \frac{\chi}{a(\chi)} \int_{\chi}^{\chi_{\rm H}} n^i(\chi')\frac{{\rm d}z}{{\rm d} \chi'}\frac{\chi' - \chi}{\chi'}{\rm d}\chi' \, .
\label{eq:q_lensing}
\end{eqnarray}
In the previous expression, $n^i(z)$ refers to the redshift distribution  in tomographic bins `$i (j)$', while $a(\chi)$ is the scale factor at co-moving distance $\chi$ from the observer.
For more details, consult \citealt{Lamman_IA_guide}.
The matter power spectrum is computed from  {\sc Halofit} \citep{Takahashi2012} in this work, however other public tools provide accurate predictions, including \eg  {\sc HMcode} \citep{HMCode2020}, the {\sc EuclidEmulator} \citep{EuclidEmulator}, the {\sc Bacco} emulator \citep{BACCOEmulator}, or the {\sc MiraTitan} emulator \citep{miraTitan}. 
Predictions for the \gtwopcf are finally computed from Eq. (\ref{eq:C_ell}) as:
\begin{eqnarray}
\xi_{\pm}^{ij}(\theta) = \frac{1}{2\pi} \int_0^{\infty} C_{\ell}^{ij} \, J_{0/4}(\ell \theta) \, \ell \, {\rm d}\ell,
\label{eq:xipm_th}
\end{eqnarray}
where $J_{0/4}(x)$ are Bessel functions of the first kind. 
In this paper, these calculations are carried out by the {\sc CosmoSIS} cosmological inference package \citep{cosmoSIS}.
The redshift distribution $n (z)$ is the source galaxy sample redshift distribution of LSST Year1 \citep{LSST-SRD}:
\begin{eqnarray}
n (z) =  z^2 \: {\rm exp} \left[-\left(\frac{z}{z_0}\right)^\alpha \right]\, ,
\label{eq:nz}
\end{eqnarray}
with pivot redshift $z_0=0.13$ and a power-law index $\alpha = 0.78$.
The distribution is normalised to provide a number density of $n_{\rm eff} = 3.0 $  galaxies per arcmin$^{2}$.
This lower than the expected number density for the first data release ($n_{\rm eff}\sim10$ galaxies arcmin$^{-2}$), but is large enough to validate our methods, and makes the calculation more tractable.
This global $n(z)$ is further split into five equi-populated tomographic bins, and smoothed with a Gaussian filter of width $\sigma = 0.05 \: (1 + z)$ to mimic the photometric selection process, shown by the solid lines in Fig. \ref{fig:Nz}.

>>>>

Shear two-point correlation functions (\gtwopcf hereafter) can be predicted with percent-level precision and are therefore an ideal quantity to validate weak lensing simulations.
In the Limber approximation\footnote{See \citet{Kilbinger17} for a comparison between the Limber approximation and the exact calculations.}, the lensing angular power spectrum $C^{ij}_{\ell}$ (in spherical harmonics space) in tomographic bins $i$ and $j$ is obtained from an integral over the three-dimensional matter power spectrum $P_{\delta}(k, z)$ (in Fourier space, and redshift $z$ dependant):
  \begin{eqnarray}
C_{\ell}^{ij} = \int_0^{\chi_{\rm H}}  \frac{q^i(\chi) \,q^j(\chi)}{\chi^2} \, P_{\delta}\, \bigg(\frac{\ell+1/2}{\chi},z(\chi)\bigg) \ {\rm d}\chi,
\label{eq:C_ell}
\end{eqnarray}
where $\ell$ is a multipole, $\chi_{\rm H}$ is the co-moving distance to the horizon, $c$ is the speed of light, $H_0$ the Hubble parameter, and the lensing kernels $q^{i}$ and $q^j$ are given by:
\begin{eqnarray}
q^{i(j)}(\chi) = \frac{3}{2}\Omega_{\rm m} \, \bigg(\frac{H_0}{c} \bigg)^2 \frac{\chi}{a(\chi)} \int_{\chi}^{\chi_{\rm H}} n^i(\chi')\frac{{\rm d}z}{{\rm d} \chi'}\frac{\chi' - \chi}{\chi'}{\rm d}\chi' \, .
\label{eq:q_lensing}
\end{eqnarray}
In the previous expression, $n^i(z)$ refers to the redshift distribution  in tomographic bins `$i (j)$', while $a(\chi)$ is the scale factor at co-moving distance $\chi$ from the observer.
For more details, consult \citealt{Lamman_IA_guide}.
The matter power spectrum is computed from  {\sc Halofit} \citep{Takahashi2012} in this work, however other public tools provide accurate predictions, including \eg {\sc HMcode} \citep{HMCode2020}, the {\sc EuclidEmulator} \citep{EuclidEmulator}, the {\sc Bacco} emulator \citep{BACCOEmulator}, or the {\sc MiraTitan} emulator \citep{miraTitan}. 
Predictions for the \gtwopcf are finally computed from Eq. (\ref{eq:C_ell}) as:
\begin{eqnarray}
\xi_{\pm}^{ij}(\theta) = \frac{1}{2\pi} \int_0^{\infty} C_{\ell}^{ij} \, J_{0/4}(\ell \theta) \, \ell \, {\rm d}\ell,
\label{eq:xipm_th}
\end{eqnarray}
where $J_{0/4}(x)$ are Bessel functions of the first kind. 
In this paper, these calculations are carried out by the {\sc CosmoSIS} cosmological inference package \citep{cosmoSIS}.
The redshift distribution $n (z)$ is the source galaxy sample redshift distribution of LSST Year1 \citep{LSST-SRD}:
\begin{eqnarray}
n (z) =  z^2 \: {\rm exp} \left[-\left(\frac{z}{z_0}\right)^\alpha \right]\, ,
\label{eq:nz}
\end{eqnarray}
with pivot redshift $z_0=0.13$ and a power-law index $\alpha = 0.78$.
The distribution is normalised to provide a number density of $n_{\rm eff} = 3.0 $  galaxies per arcmin$^{2}$.
This is a bit lower than the expected number density for the first data release ($n_{\rm eff}\sim10$ galaxies arcmin$^{-2}$), but is large enough to validate our methods, and makes the calculation more tractable.
This global $n(z)$ is further split into five equi-populated tomographic bins, and smoothed with a Gaussian filter of width $\sigma = 0.05 \: (1 + z)$ to mimic the photometric selection process, shown by the solid lines in Fig. \ref{fig:Nz}.
